---
title: "covid19api_vignette"
author: "Victoria Seng"
date: "9/21/2021"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Required Packages (will update if additional are needed)
*  `tidyverse`
*  `httr`
*  `jsonlite`

# Functions
*  `getSummary()`
*  `getCountries()`
*  `returnSlug()`
*  `getDayOneLive()`

```{r libs, echo = FALSE, message = FALSE, results = "hide"}
library(tidyverse)
library(httr)
library(jsonlite)
```

## `getSummary()`

Grabs the daily summary, which shows total and new cases by country and case type

```{r define_getSummary}

getSummary <- function(){
  resp <- GET("https://api.covid19api.com/summary")
  summary <- as_tibble(fromJSON(rawToChar(resp$content)))
  return(summary)
}

```

## `getCountries()`

Grabs all country names, slugs, and abbreviations. It is used inside all other functions where the user inputs `country` as an argument.

```{r define_getCountries}

getCountries <- function(){
  resp <- GET("https://api.covid19api.com/countries")
  countries <- as_tibble(fromJSON(rawToChar(resp$content)))
  return(countries)
}

```

## `returnSlug()`

Takes a quoted string and returns the country slug (if available) as a character string. Argument can be the name of the country, the slug, or the ISO2 abbreviation. Throws a specific error if no match is found.

TO DO: I might try to make this not case sensitive in the future

```{r returnSlug}
returnSlug <- function(cntry){
  countries <- getCountries() ## grab all the countries as a tibble
  
  slug <- countries %>%
    filter(
        case_when(
          cntry %in% Country ~ cntry == Country,
          cntry %in% ISO2    ~ cntry == ISO2,
          cntry %in% Slug    ~ cntry == Slug
      ) ## figure out which column the argument is referencing, then filter to the correct row 
    ) %>% 
      select(Slug) ## grab the slug
  
  if(nrow(slug) == 0) stop(
    paste0(
      cntry, " is not a country name, slug, or ISO2 valus supported by the api"
    ) ## if there is no match, the resulting tibble will have 0 rows. Throw an error if this happens
  )
  
  return(unname(unlist(slug))) ## turn the slug into a character string and return it
}

```

## `getDayOneLive()`

This endpoint returns all cases of the specified case type for the specified country from its first recorded case to present (updated every 10 minutes.)

Its arguments are `country` and `caseType`.

TO DO: figure out if there is an all option to this endpoint. If there is, I'll need to have these be unnamed args and add logic to accommodate.

```{r define_getDayOneLive}
getDayOneLive <- function(cntry, caseType){
  params <- vector()
  
  params[1] <- returnSlug(cntry)
  params[2] <- caseType
  
  resp <- GET(
    paste0(
      "https://api.covid19api.com/dayone/country/",params[1],"/status/",params[2],"/live"
    )
  )
  dayOneLive <- as_tibble(fromJSON(rawToChar(resp$content)))
  return(dayOneLive)
}
```


*  By Country Live params = country, case type, from
*  Live by Country and Status params = country, case type, from
*  Live by Country All Status
*  Live by Country And Status After Date
*  Will need to allow for country name, slug, or abbrev to work in args (maybe try to find partial matches too?)
*  Will need to allow for various date and datetime formats to work as arguments
*  Decide if I want to do a function per endpoint or have the endpoints passed to the function as arguments and have logic inside the function. 
  *  Looking at the examples, I think I'll try 1+ functions per endpoint and a wrapper function around them.


# Exploratory Data Analysis (EDA)

# Wrap Up