---
title: "covid19api_vignette"
author: "Victoria Seng"
date: "9/21/2021"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Required Packages (will update if additional are needed)
*  `tidyverse`
*  `httr`
*  `jsonlite`

# Functions

right now, this is just the spot where I'm testing out calling the api. Will replace this with a function later.

```{r testCall, echo = TRUE}
library(tidyverse)
library(httr)
library(jsonlite)

resp <- GET("https://api.covid19api.com/countries")
str(resp) #shows that the data is in content as a raw vector like in hw4

resPretty <- as_tibble(fromJSON(rawToChar(resp$content)))
head(resPretty, 10)
```

To do: 

*  Decide which api endpoints to support
  *  Countries (will be necessary anyway as a lookup for function args)
  *  Summary
  *  Day One Live
  *  By Country Live
  *  Live by Country and Status
  *  Live by Country All Status
  *  Live by Country And Status After Date
*  Will need to allow for country name, slug, or abbrev to work in args (maybe try to find partial matches too?)
*  Will need to allow for various date and datetime formats to work as arguments
*  Decide if I want to do a function per endpoint or have the endpoints passed to the function as arguments and have logic inside the function. 
  *  Looking at the examples, I think I'll try 1+ functions per endpoint and a wrapper function around them.


# Exploratory Data Analysis (EDA)

# Wrap Up